---
title: "HW1 HGEN 48800 Lin_Yu"
author: "Lin_Yu"
date: "2024-03-31"
fontsize: 12pt
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```




### Problem 1
(a)
i. Implement the linear search algorithm
```{r} 
# Implement the linear search algorithm, which scans through a given sequence, looking for an element v
linear_search <- function(arr, target) {
  n <- length(arr)
  for (i in 1:n) {
    if (arr[i] == target) {
      return(i)  # Return the index where the target is found
    }
  }
  return("NULL")  # If target is not found, return "NULL"
}
```

```{r} 
# simple test:
arr <- c(1, 5, 3)
target <- 5
result <- linear_search(arr, target)
if (result == "NULL") {
  cat("Target", target, "not found in array.")
} else {
  cat("Target", target, "found at index", result)
}

```
ii.Loop invariant to prove the correctness of our algorithm:

-Base case:
At iteration 0, the algorithm checks if A[1] equals v. If it does, the algorithm returns the index 0. Otherwise, it returns "Null".

-Induction Step:
Assuming the algorithm works correctly for the first (k-1) iterations without finding v, at the kth iteration, it checks if A[k] equals v. If v is not found among the first k elements, it is not in the array. If v is found at index k, the algorithm returns k. 

Therefore, the algorithm operates correctly for all k iterations.

(b) 
i.The average number of elements needed to be checked:
- If the one we are searching for is equally likely to be any element in sequence A,  
let X be the index of the element we search for, Y be the number of elements we checked. Then we know Y=X
The average number of elements needed to be checked 
$$
E(Y)=\sum_{i=1}^{n}p(X=i)X=\frac{1}{n}\sum_{i=1}^{n}i=\frac{n+1}{2}
$$
ii.The number of elements needed to be checked in worset case:
The worst case: what we are looking for is $a_{n}$, we need n iterations.

iii.The average-case running time and the worst case should be similar, as O(n).

(c)
- Sort the sequence A
```{r}
Insertion_sort <- function(arr) {
  n <- length(arr)
  for (j in 2:n) {
    key <- arr[j]
    i <- j-1
    while (i>0 && arr[i] >key ) {
      arr[i+1] <- arr[i]
      i=i-1
    } 
      arr[i+1] <- key  
  }
  return(arr)
  }
```

-implement binary search
```{r}
binary_search <- function(arr,target){
  n <- length(arr)
  a=1
  b=n
  while (a<=b){
    mid <-  floor((b+a)/2)
    if (target == arr[mid]){
      return(mid)
    }else if (arr[mid] < target) {
      a=mid+1
    }else{
      b=mid-1
    }
  }
  return("NULL")  # If target is not found, return "NULL"
    
}

```


```{r}
# test:
arr <- c(1, 3, 13, 5, 15, 17, 19)
arr <- Insertion_sort(arr)
print(arr)
target <- 13
result <- binary_search(arr, target)
if (result == "NULL") {
  cat("Target", target, "not found in array.")
} else {
  cat("Target", target, "found at index", result, "in the sorted sequence of A")
}
```

(d)

The worst-case should be when target element is not in the array, but lies in the range of $[a_{1},a_{n}]$. 
The worst-case time complexity of binary search is O(log n), where n is the number of elements in the array

Reasoning: 

- As the algorithm will continue to divide the array in half until it reaches a point where the subarray to be searched is empty,
let k be the number of iterations needed,
$$n/2^k=1$$
$$ k = \log_2{n} $$

### Problem 2 

It is in the handwritten section :)

### Problem 3 

Merge-sort Algorithm

i. implement the merge-sort algorithm for the sorting problem:
```{r}

merge <- function(arr, p, q, r) {
  n1 <- q - p + 1
  n2 <- r - q
  L <- arr[p:q]
  R <- arr[(q + 1):r]
  inf <- max(arr) * 100

  L <- c(L, inf)
  R <- c(R, inf)

  i <- 1
  j <- 1
  for (k in p:r) {
    if (L[i] <= R[j]) {
      arr[k] <- L[i]
      i <- i + 1
    } else {
      arr[k] <- R[j]
      j <- j + 1
    }
  }
  return(arr)
}
```

```{r}
merge_sort <- function(arr, p, r) {
  if (p < r) {
    q <- floor((p + r) / 2)
    arr=merge_sort(arr, p, q)
    arr=merge_sort(arr, q + 1, r)
    merge(arr, p, q, r)
  }else{
    arr
  }
}

# Test the merge-sort function
arr1 <- c( 4,3,1,3,8,7,2)
sort_arr=merge_sort(arr1,1,length(arr1))
print(sort_arr)
```

### Problem 4 

i.pseudo code: 



```r

Function hash_sequence(tuple):
//Assign 3-tuple a value (hash function)
    A <- 0
    C <- 1
    G <- 2
    T <- 3
    hash_value <- 0
    hash_value += value_of(tuple[0]) * (4 ** 2)
    hash_value += value_of(tuple[1]) * (4 ** 1)
    hash_value += value_of(tuple[2]) * (4 ** 0)
    return hash_value

Function positions_table(sequence1, sequence2, ..., sequencek):
// Generate the position_table using the provided sequences as our dataset

    positions_table <- empty position tables
    for j from 1 to k: 
    // Iteration over all sequences j, j=1,2,..k
        for i from 1 to length(sequence j) - 2: 
        // check all the 3-tuple and store them into the position table
            tuple <- substring of sequencej from index i to i + 2
            hash_value <- hash_sequence(tuple)  
            // Compute hash value for this tuple
            if (hash_value is in the positions_table):
               Append index (j, i) to positions_table[hash_value] 
            else:
               create a new row in the positions_table
    return positions_table

Function search_hash_table(target_sequence, positions_table):
// This function generates a hash_table based on the target sequence and
// determines if the target sequence is found.

    m <- length(target_sequence) / 3
    hash_table <- empty data frame
    for s from 0 to m - 1:
        tuple <- substring of target_sequence from index 3 * s + 1 to 3 * s + 3
        hash_value <- hash_sequence(tuple)
        for index(j, i) stored in positions_table[hash_value]:
            append tuple (j, i - 3 * s, i) as Sequence_index,target_start_Position 
            and start_Position to hash_table
    M <- sorted hash_table
    for each m consecutive tuples(rows) in M:
        if all tuples(rows) have the same first index:
           if all rows have the same second index:
              return the first tuple of the m consecutive tuples(rows)
    return "Pattern not found"
```    


ii. Implementing the pseudo code
```{r}
# Assign 3-tuple a value (hash function)
hash_sequence <- function(tuple) {
  nucleotides <- c("A" = 0, "C" = 1, "G" = 2, "T" = 3)
  hash_value <- 0
  for (i in 1:3) {
    hash_value <- hash_value + nucleotides[substring(tuple, i, i)] * (4 ** (3 - i))
    
  }
  return(as.integer(hash_value))
}

#Test if the hash_sequence works
hash_sequence("CGA")
```


```{r}
# Create an empty data frame with appropriate column names
positions_table <- data.frame(Hash_Value = numeric(), 
                              Substring = character(), 
                              Position_Index = I(list()), 
                              stringsAsFactors = FALSE)

# Generate the position_table using the provided sequences as our dataset
build_positions_table <- function(...) {
  sequences <- list(...)
  
  for (j in 1:length(sequences)) {
    for (i in 1:(nchar(sequences[[j]]) - 2)) {
      substring <- substr(sequences[[j]], i, i + 2)  # Extract substring from the sequence
      hash_value <- hash_sequence(substring)
      
      # Check if the hash value already exists in the positions table
      if (hash_value %in% positions_table$Hash_Value) {
        # If yes, update the corresponding row
        idx <- which(positions_table$Hash_Value == hash_value)
        positions_table$Position_Index[[idx]] <- c(positions_table$Position_Index[[idx]], paste("(", j, ",", i, ")", sep = ""))
      } else {
        # If no, create a new row
        new_row <- data.frame(Hash_Value = hash_value, 
                              Substring = substring,
                              Position_Index = I(list(paste("(", j, ",", i, ")", sep = ""))))
        positions_table <- rbind(positions_table, new_row)
      }
    }
  }
  
  return(positions_table)
}
# Example usage
sequence1 <- "GCTGCT"
positions_table <- build_positions_table(sequence1)

```


```{r}
#convert string "(j,i)" to numerical list (j,i)
translate_string <- function(string) {
  # Remove parentheses and split the string by comma
  components <- strsplit(gsub("[()]", "", string), ",")[[1]]
  
  # Convert components to numeric and store as a list
  numeric_pair <- list(j = as.numeric(components[1]), i = as.numeric(components[2]))
  
  return(numeric_pair)
}

x=translate_string("(3,6)")

#Create the hash_table based on target_sequence and positions_table
hash_table <- function(target_sequence, positions_table) {
  m <- nchar(target_sequence) / 3
  #create an empty data frame for hash_table
  M <- data.frame(Sequence_index = numeric(),
                  target_start_Position = numeric(),
                  start_Position = numeric(),
                  stringsAsFactors = FALSE)
                            
  # Loop through substrings of target_sequence
  for (s in 0:(m - 1)) {
    start <- 3 * s + 1
    end <- 3 * s + 3
    target_tuple <- substr(target_sequence, start, end)# extract(s+1)th 3-tuples in the target_sequence
    hash_value <- hash_sequence(target_tuple)
    
    # Check if hash_value of target_tuple exists in positions_table
    if (hash_value %in% positions_table$Hash_Value) {
      idx <- which(positions_table$Hash_Value == hash_value)
      position_indexes=positions_table$Position_Index[[idx]]
      for (indexes in position_indexes) {
        index <- translate_string(indexes)
        # Append (j, i - 3 * s, i) to hash_table
        new_row <- data.frame(Sequence_index = as.numeric(index[1]),
                              target_start_Position = as.numeric(index[2]) - 3 * s,
                              start_Position = as.numeric(index[2])) 
                              # the place where the first letter in target tuple occurs in sequence i
        M <- rbind(M, new_row) # Add new row to M
      }
    }
  }
  return(M)
}

M=hash_table('GCT', positions_table)
```



```{r}
# This function generates the hash_table based on target sequence and 
# determines if it founds the target sequence
search_hash_table <- function(target_sequence,positions_table) {
  m <- nchar(target_sequence) / 3
  M <- hash_table(target_sequence, positions_table)
  sorted_M <- M[order(M$Sequence_index, M$target_start_Position, M$start_Position), ]
  found <- 0
  # Check for m consecutive tuples in M
  for (i in 1:(nrow(sorted_M) - m + 1)) {
    #Check if the first and second indexes of the tuple are the same
    if (all(sorted_M$Sequence_index[i:(i + m - 1)] == sorted_M$Sequence_index[i]) &&
        all(sorted_M$target_start_Position[i:(i + m - 1)] == sorted_M$target_start_Position[i])) {
      cat("Our target occurs at sequence", sorted_M$Sequence_index[i], "starting from position", sorted_M$target_start_Position[i], "\n")
      found <-found+1
      #return(invisible(NULL))  # Exit the function if pattern is found
    }
  }
  if (found==0){
    cat("Pattern not found\n")
  }
  # Print "Pattern not found" if no match is found
}


```

```{r}
target="TAGCTAGCT"
S1 = "GCTGCTGCTGCTAAACGTTTGGGGCAGTCGAT"
S2 = "GGTGCTCCAAGCTTTTGAGTCTGCTAGTGTCAACCCT" 
S3 = "GTGGGCCCCCTAGCTAGCTAGCTGGGGCAC"
S4 = "TGTCGCTGGCTGGACTGCTGATCGTAGTAG"
positions_table <- build_positions_table(S1,S2,S3,S4)
search_hash_table(target,positions_table)
```
   
    
    


